---
title: "assignment_2.1"
author: "Charles Hendrickson"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Assignment 2.1 – dataRetrieval
1) Add the Santa Paula Creek gauge to exercise we did in class
2) What is the best way to query multiple sites using this R package?
3) Create a plot showing any stream flow response to the rain of last night (10/04). Determine at what time did the stream flow peaked (if any) at the two locations?



What we did in class (for reference):
```{r}
library(dataRetrieval)
library(tidyverse)
library(kableExtra)
```

# Exercise 1
# Start a new Markdown document to plot the discharge time-series for the Ventura River from 2019-10-01 to 2020-10-05
```{r}
# Ventura River discharge time-series:
siteNumber <- "11118500"
parameterCd <- "00060"  # Discharge
startDate <- "2019-10-01"  
endDate <- "2020-10-05" 

discharge <- readNWISdv(siteNumber, 
                    parameterCd, startDate, endDate)

```


```{r}
discharge <- discharge %>%
  rename(discharge = X_00060_00003)
```


```{r}
ggplot(data = discharge, aes(x = Date, y = discharge)) +
  geom_line() +
  labs(title = "USGS Discharge for the Ventura River",
       x = "Date",
       y = "Discharge (cubic feet per second)") +
  theme_classic()
```


# Now add the Santa Paula Creek gauge to exercise we did in class:
# Question: What is the best way to query multiple sites using this R package?

# Answer: I think the best way to query multiple sites using this R package is to combine both sites into a vector, which puts both sites into the same data frame. 

Santa Paula Creek and Ventura River discharge time-series for October 4-5th:
```{r}
# Santa Paula Creek discharge time-series:
SP_and_V_discharge <- readNWISdata(sites= c("11113500","11118500"), service="iv", 
                   parameterCd="00060", 
                   startDate="2021-10-04T00:00Z", endDate="2021-10-05T11:59Z",
                   tz = "America/Los_Angeles")
```

```{r}
# Rename the column `X_00060_00000` as `discharge`
SP_and_V_discharge <- SP_and_V_discharge %>% 
  rename(discharge = X_00060_00000)
```

# Create a plot showing any stream flow response to the rain of last night (10/04). Determine at what time did the stream flow peaked (if any) at the two locations?

```{r}
library(ggplot2)

ggplot(data = SP_and_V_discharge, aes(x = dateTime, y = discharge)) +
  geom_line(aes(color = site_no)) +
  labs(title = "USGS Discharge for Santa Paula Creek and Ventura River 
(10/04/2021 - 10/05/2021)",
       x = "Time",
       y = "Discharge (cubic feet per second)") +
  scale_color_discrete(name = "Site", labels = c("Santa Paula Creek (11113500)", "Ventura River (11118500)")) +
  theme_classic()
```

Determine at what time did the stream flow peaked (if any) at the two locations?
```{r}
# Create separate data frames for the two sites 
santa_paula_creek <- SP_and_V_discharge %>% 
  filter(site_no == 11113500)

ventura_river <- SP_and_V_discharge %>% 
  filter(site_no == 11118500)

```


```{r}
# Find what the max discharge was for the two sites 
max_flow_santa_paula_creek <- max(santa_paula_creek$discharge)
max_flow_santa_paula_creek

max_flow_ventura_river <- max(ventura_river$discharge)
max_flow_ventura_river
```

```{r}
# Filter for the time when the max discharge occurred 
time_max_flow_santa_paula <- santa_paula_creek %>%
  filter(discharge == max_flow_santa_paula_creek) %>% 
  select(dateTime, discharge)

kable(time_max_flow_santa_paula)

time_max_flow_ventura <- ventura_river %>%
  filter(discharge == max_flow_ventura_river) %>% 
  select(dateTime, discharge)

kable(time_max_flow_ventura)

```












