---
title: 'exercise 2'
author: "Charles Hendrickson"
date: "10/6/2021"
output: html_document
---

Let’s determine what percentage of Alaskan household are speaking only English!

The data: https://doi.org/10.5063/F1N58JPP

1) Read the metadata
2) Download the data household_language.csv using metajam
3) Read the data into R using metajam
4) Write a piece of code that will compute the percentage of Alaskan household speaking only English for the year 2009 to 2015
5) Create a plot to visualize this data



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Install metjam
# install.packages("metajam")
```

```{r}
# Load packages 
library(metajam)
library(tidyverse)
library(here)
```

```{r}
# set inputs
data_obj <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
path <- "~/Documents/eds213/week_2/data_2"
# download data and metadata
# download_d1_data(data_obj, path)
```

```{r}
# Read the data 
my_data <- metajam::read_d1_files("~/Documents/eds213/week_2/data_2/doi_10.5063_F1CJ8BPH__household_language__csv")

```

Write a piece of code that will compute the percentage of Alaskan household speaking only English for the year 2009 to 2015
```{r}
# Converting my_data from a list to a data frame 
my_data <- my_data$data

library(dplyr)
library(tidyverse)

years_to_include <- c(2009:2015)

percent_english_only <- my_data %>%
  select(Year, speak_only_english, total) %>%
  filter(Year %in% years_to_include) %>% 
  group_by(Year) %>% 
  summarize(sum_speak_only_english = sum(speak_only_english),
            sum_total = sum(total)) %>% 
  mutate(Percent = 100 * (sum_speak_only_english / sum_total))

percent_english_only

```

Create a plot to visualize this data
```{r}
ggplot(data = percent_english_only, aes(x = Year, y = Percent)) +
  geom_line() +
  labs(title = "Percentage of Alaskan Households Speaking Only English (2009-2015)",
       x = "Year",
       y = "Percent Speaking Only English (%)") +
  theme_classic()
```




